---
title: "R Notebook"
output: html_notebook
---

# Load Data

```{r}
library(rpart)
library(randomForest)
library(party)
library(caret)
library(tidyverse)

set.seed(415)

training.data <- read_csv('../data/train.csv') %>%
  mutate(Set = 'train')

testing.data <- read_csv('../data/test.csv') %>%
  mutate(Survived = 1,
         Set = 'test')

combined.data <- training.data %>%
  union(testing.data)

combined.data
```

# Perform Imputations and Transformations

```{r}
combined.data <- combined.data %>%
  separate(Name, c('Surname','Title','First'), sep='[.,]') %>%
  mutate(Group = factor(Survived, levels=0:1, labels=c('Perished','Survived')),
         FamilySize = SibSp + Parch + 1,
         Members = FamilySize,
         Title = str_trim(Title),
         Surname = str_trim(Surname),
         FamilyName = Surname,
         First = str_trim(First),
         Title = recode(Title,
                        'Mr' = 'Mr',
                        'Miss' = 'Miss',
                        'Mrs' = 'Mrs',
                        'Master' = 'Master',
                        'Ms' = 'Miss',
                        'Mlle' = 'Miss',
                        'Mme' = 'Mrs',
                        .default = 'Titled'),
         Embarked = ifelse(is.na(Embarked), 'S', Embarked),
         Fare = ifelse(is.na(Fare), median(Fare, na.rm = TRUE), Fare)) %>%
  unite('FamilyID',Members,FamilyName) %>%
  mutate(FamilyID2 = ifelse(FamilySize <= 3, 'Small', FamilyID),
         Pclass = factor(Pclass),
         Surname = factor(Surname),
         Title = factor(Title),
         Embarked = factor(Embarked),
         FamilyID = factor(FamilyID),
         FamilyID2 = factor(FamilyID2),
         Sex = factor(Sex))

AgeFit <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title + FamilySize,
                data = combined.data[!is.na(combined.data$Age),],
                method = 'anova')

combined.data$Age[is.na(combined.data$Age)] <- predict(AgeFit, combined.data[is.na(combined.data$Age),])

# Housekeeping: clear unused objects from memory
rm(AgeFit)

combined.data
```

# Re-create Training Data

```{r}
training.data <- combined.data %>%
  filter(Set == 'train') %>%
  select(Survived,
         Group,
         Pclass,
         Sex,
         Age,
         SibSp,
         Parch,
         Fare,
         Embarked,
         Title,
         FamilySize,
         FamilyID,
         FamilyID2)

training.data
```

# Re-create Test Data

```{r}
testing.data <- combined.data %>%
  filter(Set == 'train') %>%
  select(Survived,
         Group,
         Pclass,
         Sex,
         Age,
         SibSp,
         Parch,
         Fare,
         Embarked,
         Title,
         FamilySize,
         FamilyID,
         FamilyID2)

testing.data
```

# Output Data Files

```{r}
system.time({
  write_csv(combined.data,
          '../data/combined.csv',
          na='')
})

system.time({
  write_csv(training.data,
          '../data/clean_training.csv',
          na='')
})

system.time({
  write_csv(testing.data,
          '../data/clean_testing.csv',
          na='')
})
```

# Random Forest

## Summary

```{r}
fit.1 <- randomForest(Group ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize,
                    data = training.data,
                    importance = TRUE,
                    ntree = 2000)

Prediction.1 <- predict(fit.1, training.data)

summary(fit.1)
```

## Confusion Matrix

```{r}
table(training.data$Group, Prediction.1)

```

## Variable Importance

```{r}
varImpPlot(fit.1)
```

# Support Vector Machine

## Untuned Model

### Summary

```{r}
fit.2 <- train(Group ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize,
               data = training.data,
               method = 'svmRadial',
               trControl = trainControl(method='repeatedcv', number=10, repeats=3),
               preProcess = c('center','scale'),
               tuneLength = 10)

Prediction.2 <- predict(fit.2, training.data)

fit.2
```

### Confusion Matrix

```{r}
confusionMatrix(Prediction.2, training.data$Group)
```

### Model Parameters

```{r}
plot(fit.2)
```

## Tuned Model

### Summary

```{r}
my.grid <- expand.grid(sigma=c(0.01,0.02,0.025,0.03,0.04,0.05, 0.06, 0.07,0.08, 0.09, 0.1, 0.25, 0.5, 0.75,0.8,0.9),
                       C=c(0.01, 0.05, 0.1, 0.15, 0.25, 0.5, 0.75,1, 1.5, 2,5))
fit.3 <- train(Group ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize,
               data = training.data,
               method = 'svmRadial',
               trControl = trainControl(method='repeatedcv', number=10, repeats=3),
               preProcess = c('center','scale'),
               tuneGrid = my.grid,
               tuneLength = 10)

Prediction.3 <- predict(fit.3, training.data)

fit.3
```

### Confusion Matrix

```{r}
confusionMatrix(Prediction.3, training.data$Group)
```

### Tuning Parameters

```{r}
plot(fit.3)
```

### Tuning Grid

```{r}
as_data_frame(fit.3$results) %>%
  mutate(sigma = factor(sigma),
         C = factor(C)) %>%
  ggplot(aes(sigma, C)) +
    geom_tile(aes(fill=Accuracy), color='white') +
    scale_fill_gradient(low='lightgrey', high='steelblue') +
    ylab('Cost') +
    xlab('sigma') +
    theme(legend.title = element_text(size = 10),
          legend.text = element_text(size = 12),
          plot.title = element_text(size = 16),
          axis.title = element_text(size = 14, face = 'bold'),
          axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(fill = 'Accuracy')

```

# Various Plots and Outputs

## Age By Group Box Plot

```{r}
training.data %>%
  mutate(Group = factor(Survived, levels=0:1, labels=c('Perished','Survived'))) %>%
  ggplot(aes(x=Group,
             y=Age)) + geom_boxplot(na.rm=TRUE)
```

## Age by Group Frequency Polygon

```{r}
training.data %>%
  # filter(Survived==1) %>%
  mutate(Group = factor(Survived, levels=0:1, labels=c('Perished','Survived'))) %>%
  ggplot(aes(Age, color=Group)) + geom_freqpoly(na.rm=TRUE)
```

## Titles

```{r}
training.data %>%
  group_by(Title) %>%
  summarise(Count = n(),
            Survivors = sum(Survived),
            Casualties = Count - Survivors) %>%
  ungroup() %>%
  mutate(Percentage_Surv = Survivors / Count,
         Percentage_Per = Casualties / Count) %>%
  ggplot(aes(x=reorder(Title, Percentage_Surv), y=Percentage_Surv)) +
     geom_col(fill='darkblue') +
     scale_x_discrete(name="Title") +
     scale_y_continuous(name="Percentage",
                        limits = c(0, 1),
                        labels = scales::percent) +
     ggtitle('Percentage Survived By Title') +
     coord_flip()

```




